Portfolios of Subgraph Isomorphism Algorithms

### 2015-12-16 paper/paper.tex
Christine's changes

In Section~\ref{sec:algs}, we experimentally compare eight state-of-the-art
algorithms. We introduce a large benchmark set composed of 5725 instances
composed of twelve classes. Ten of these classes have been introduced in the
experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}. We
evaluate the algorithms, of which two are novel, on this benchmark set, and
show that they have very complementary performance. In particular, we show that
depending on the CPU time limit, different algorithms achieve the best
performance on the entire benchmark set.

@@

Given a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$, the \emph{subgraph
isomorphism problem} consists of deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
More formally, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
different target node to each pattern node, and preserves pattern edges, i.e.\ $\forall (u,u')
\in E_p, (f(u),f(u')) \in E_t$.
Note that the subgraph is not necessarily induced, so that two pattern nodes not linked by
an edge may be mapped to two target nodes which are linked by an edge. This problem is also called
\emph{subgraph monomorphism} or \emph{subgraph matching} in the literature.

@@

The simplest form of filtering is simply to propagate difference constraints (which ensure that the
matching is injective) and edge constraints (which ensure that the matching preserves pattern
edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes
every candidate pair $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or
$(u,u')$ is a pattern edge but $(v,v')$ is not a target edge (edge constraint). This simple
filtering (called \emph{Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for
difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for
example, in McGregor's algorithm~\cite{mcgregor79} and in \VFtwo{}~\cite{Cordella:2004}.

R\'egin~\cite{regin} introduced a stronger filtering for difference constraints, which ensures that
all pattern nodes can be matched with different target nodes, all together. This filtering (called
\emph{All-Different Generalized Arc Consistency}) removes more candidate pairs than when each
difference constraint is propagated separately which Forward-Checking. However, it is also more time
consuming as requires ${\cal O}(n_p^2\cdot n_t^2)$ time.

@@

We have selected a set of algorithms with complementary performance. These
algorithms will compose our portfolio in the per-instance selection approach
described in Section~\ref{sec:algsel}.
\begin{itemize}
\item \VFtwo{}~\cite{Cordella:2004} performs weak filtering that it is especially fast on
    trivially satisfiable instances;
\item \LAD{}~\cite{Solnon:2010} combines two strong but expensive filtering techniques
    (All-Different Generalized Arc Consistency and Locally All-Different);
\item \Glasgow{}~\cite{McCreesh:2015} does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic All-Different propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}

@@

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs~\cite{constraints10}.
\item Classes 2 and 3 contain various kinds of graphs selected by Larrosa and Valiente~\cite{LV02}:
    class 2 contains small instances generated from the 50 first graphs of class
    1 and class 3 contains
    larger instances with pattern graphs from class 2 and target graphs from
    the next 50 graphs.
\item Classes 4 to 8 contain randomly generated graphs from a database of graphs commonly
    used for benchmarking subgraph isomorphism algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for classes 4 and 5, regular meshes
    for classes 6 and 7, and random graphs with uniform edge probabilities for class 8. All of these
    instances are satisfiable.
\item Classes 9 and 10 contain instances from segmented images~\cite{pr15,cviu11}.
\item Class 11 contains instances from meshes modeling 3D objects~\cite{cviu11}.
\item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
    phase transition---these instances are expected to be particularly challenging, despite their
    small size.
\end{itemize}

@@

Furthermore, the virtual best solver (VBS), which considers the best algorithm
for each instance separately, obtains much better results, showing us that the
algorithms have complementary performance. The difference between VBS and single
best is particularly pronounced for CPU time limits less than \SI{1000}{\ms}. In
many applications, it is important to have the fastest possible algorithm even
if the absolute differences in CPU time are small. For example, in pattern
recognition~\cite{pr15,cviu11} and chemical~\cite{Giugno:2013} applications, we
often have to solve the subgraph isomorphism problem repeatedly for a very large
number of graphs (in order to find a pattern image or molecule in a large
database of target images or compounds, for example), so having an algorithm
that is able to solve an instance in \SI{100}{\ms} instead of \SI{1000}{\ms}
makes a big difference.  Therefore, it is important to select the best algorithm
for each instance, even if the instance is an easy one.

@@

Experimental results reported in Section~\ref{expComp} have shown us that \IncompleteLAD is very fast
(\SI{7}{\ms} on average) and is able to solve 1919 instances from our benchmark
set very quickly. Therefore, we first run
\IncompleteLAD: if unsatisfiability is detected, we do not need to process it further.
Otherwise, we record the number of candidate pairs removed and as a percentage
of all pairs and the minimum and maximum values on a per-variable basis. The
recorded values are then used as features for algorithm selection.

@@

If presolving does not give us a solution, we extract additional features. For both the pattern and
the target graph, we consider some basic graph properties, which can be computed
very quickly,
\begin{itemize}
    \item the number of vertices and edges;
    \item the density---we expect that some kinds of filtering might be expensive and ineffective on
        dense graphs;
    \item how many loops (self-adjacent vertices) the graph contains---since loops must be mapped to
        loops, this could have a strong effect on how easy an instance is;
    \item the mean and maximum degrees, and whether or not every vertex has the same degree (the
        degree-based invariants used by \LAD and \Glasgow do nothing at the top of search if every
        vertex has the same degree);
    \item whether or not the graph is connected;
    \item the mean and maximum distances between all pairs of vertices (if nearly all vertices are
        close together, path-based reasoning is likely to be ineffective) and
        the proportion of vertex pairs which are at least 2, 3 and 4 apart.
\end{itemize}

@@

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines. The virtual best solver is the oracle predictor that, for each instance, chooses the best
solver from our portfolio. This is the upper bound of what an algorithm selection approach can
achieve. The single best solver is the one solver from the portfolio that has the overall best
performance across the entire set of instances, at the CPU time limit of \SI{e8}{\ms},
i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
algorithm selection approach.
We are able to solve 34 more instances than the single best solver within the timeout, with only an
additional 13 to the virtual best. In terms of average performance, we are able to close almost
\SI{70}{\percent} of the gap between the single best and the virtual best solver.

@@

\begin{table}[p]
    \centering\setlength{\tabcolsep}{1em}
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 2219 & 5822809\\
\LLAMA & 610835 & 2206 & 6435301\\
\GlasgowTwo & 1965158 & 2172 & 7787967\\
   \bottomrule \\
\end{tabular}
\caption{Algorithm selection performance on the reduced set of 2336
instances. MCP is the misclassification penalty; that is, the additional time
required to solve an instance because of choosing solvers that perform
worse than the best.}\label{tab:res}
\end{table}

@@

Table~\ref{tab:res} shows the performance of the selection model on its own. The performance of the entire
algorithm selection system including the preprocessing is shown in
Table~\ref{tab:resfull}.
Our system is able to close more than \SI{50}{\percent} of the gap between
single and virtual best.

@@

\begin{figure}[p]
    \input{graph-full}
\caption{Number of solved instances over time for solvers, virtual best
solvers, and \LLAMA on the full set of 5725 instances. The
actual solvers are shown as dashed lines.}
\label{fig:portfolio-ecdf-full}
\end{figure}

@@

The problem of identifying subgraph isomorphisms is a hard computation problem
that has many applications in diverse areas. In this paper, we presented a
portfolio of six algorithms from the literature and a novel variant of the \LAD
algorithm. We introduced a
set of novel features to characterise subgraph isomorphism problems and
leveraged them to select the most appropriate algorithm from the portfolio for
each instance.

We demonstrated that our algorithm selection approach achieves substantial
performance improvements over the single algorithm that has the best performance
on our benchmark set. We showed that combining an algorithm selection approach
with a novel incomplete variant of \LAD that is able to detect inconsistencies
and a presolver boosts performance even further.

Mention parallel, directed, labelled in here briefly.

>>>>

In Section~\ref{sec:algs}, we experimentally compare eight state-of-the-art
algorithms. We introduce a large benchmark set composed of 5725 instances
grouped into twelve classes. Ten of these classes have been introduced in the
experimental study reported by McCreesh and Prosser~\cite{McCreesh:2015}. We
evaluate the algorithms on this benchmark set, and show that they have very
complementary performance. In particular, we show that depending on the CPU
time limit, different algorithms achieve the best performance on the entire
benchmark set.

@@

Given a pattern graph $G_p=(N_p,E_p)$ and a target graph $G_t=(N_t,E_t)$, the \emph{subgraph
isomorphism problem} consists of deciding whether $G_p$ is isomorphic to some subgraph of $G_t$.
More formally, the goal is to find an injective matching $f: N_p\rightarrow N_t$, that associates a
different target node to each pattern node, and preserves pattern edges, i.e.\ $\forall (u,u')
\in E_p, (f(u),f(u')) \in E_t$.
Note that the subgraph is not necessarily induced, so that two pattern nodes not linked by
an edge may be mapped to two target nodes which are linked by an edge.

@@

The simplest form of filtering is to propagate difference constraints (which ensure that the
matching is injective) and edge constraints (which ensure that the matching preserves pattern
edges): each time a pattern node $u\in N_p$ is matched with a target node $v\in N_t$, one removes
every candidate pair $(u',v')\in N_p\times N_t$ such that either $v'=v$ (difference constraint) or
$(u,u')$ is a pattern edge but $(v,v')$ is not a target edge (edge constraint). This simple
filtering (called \emph{Forward-Checking}) is very fast to achieve: in ${\cal O}(n_p)$ for
difference constraints, and in ${\cal O}(d_p\cdot n_t)$ for edge constraints. It is used, for
example, in McGregor's algorithm~\cite{mcgregor79} and in \VFtwo{}~\cite{Cordella:2004}.

R\'egin~\cite{regin} introduced a stronger filtering for difference constraints, which ensures that
all pattern nodes can be matched with different target nodes, all together. This filtering (called
\emph{All-Different Generalized Arc Consistency}) removes more candidate pairs than when each
difference constraint is propagated separately which Forward-Checking. However, it is also more time
consuming as it requires ${\cal O}(n_p^2\cdot n_t^2)$ time.

@@

We have selected a set of algorithms with complementary performance. These
algorithms will compose our portfolio in the per-instance selection approach
described in Section~\ref{sec:algsel}.
\begin{itemize}
\item \VFtwo{}~\cite{Cordella:2004} performs weak filtering that is especially fast on
    trivially satisfiable instances;
\item \LAD{}~\cite{Solnon:2010} combines two strong but expensive filtering techniques
    (All-Different Generalized Arc Consistency and Locally All-Different);
\item \Glasgow{}~\cite{McCreesh:2015} does expensive preprocessing based on path length
    invariant properties to generate additional constraints, followed by weaker filtering
    (forward-checking, and a heuristic All-Different propagator which can miss deletions) and
    conflict-directed backjumping during search.
\end{itemize}

@@

\begin{itemize}
\item Class 1 contains randomly generated scale-free graphs~\cite{constraints10}.
\item Classes 2 and 3 contain instances built from a database containing
    various kinds of graph gathered by Larrosa and Valiente~\cite{LV02}: class
    2 contains small instances generated from the first 50 graphs of the
    database, and class 3 contains larger instances with pattern graphs from
    the first 50 graphs of the database and target graphs from the next 50
    graphs.
\item Classes 4 to 8 contain randomly generated graphs from a database of
    graphs commonly used for benchmarking subgraph isomorphism
    algorithms~\cite{GraphDatabase1,GraphDatabase2}: bounded-degree graphs for
    classes 4 and 5, regular meshes for classes 6 and 7, and random graphs with
    uniform edge probabilities for class 8. All of these instances are
    satisfiable.
\item Classes 9 and 10 contain instances from segmented images~\cite{pr15,cviu11}.
\item Class 11 contains instances from meshes modeling 3D objects~\cite{cviu11}.
\item Class 12 contains random graph instances chosen to be close to the satisfiable-unsatisfiable
    phase transition---these instances are expected to be particularly challenging, despite their
    small size.
\end{itemize}

@@

Furthermore, the virtual best solver (VBS), which considers the best algorithm
for each instance separately, obtains much better results, showing us that the
algorithms have complementary performance. The difference between VBS and single
best is particularly pronounced for CPU time limits less than \SI{1000}{\ms}. In
many applications, it is important to have the fastest possible algorithm even
if the absolute differences in CPU time are small. For example, in pattern
recognition~\cite{pr15,cviu11} and chemical~\cite{Giugno:2013} applications, we
often have to solve the subgraph isomorphism problem repeatedly for a very large
number of graphs (in order to find a pattern image or molecule in a large
database of target images or compounds, for example), so having an algorithm
that is able to solve an instance in \SI{100}{\ms} instead of \SI{1000}{\ms}
makes a big difference.  Therefore, it is important to select the best algorithm
for each instance, even if the instance is an easy one. Furthermore, this
selection process should not penalise easy instances, i.e.\ it should not take more time than
the solution process time for these instances.

@@

Experimental results reported in Section~\ref{expComp} have shown us that \IncompleteLAD is very fast
(\SI{7}{\ms} on average) and is able to solve 1919 instances from our benchmark
set very quickly. Therefore, we first run
\IncompleteLAD: if unsatisfiability is detected, we do not need to process it further.

@@

If presolving does not give us a solution, we extract features. For both the pattern and
the target graph, we consider some basic graph properties, which can be computed
very quickly:

\begin{itemize}
    \item the number of vertices and edges;
    \item the density---we expect that some kinds of filtering (like those based upon locally
        all-different constraints) might be expensive and ineffective on
        dense graphs;
    \item how many loops (self-adjacent vertices) the graph contains---since loops must be mapped to
        loops, this could have a strong effect on how easy an instance is;
    \item the mean and maximum degrees, and whether or not every vertex has the same degree (the
        degree-based invariants used by \LAD and \Glasgow do nothing at the top of search if every
        vertex has the same degree);
    \item whether or not the graph is connected;
    \item the mean and maximum distances between all pairs of vertices (if nearly all vertices are
        close together, path-based reasoning is likely to be ineffective) and
        the proportion of vertex pairs which are at least 2, 3 and 4 apart.
\end{itemize}

Alongside these basic features, we include information computed by \IncompleteLAD. To (try to) prove
inconsistency, \IncompleteLAD removes candidate pairs. The number of successfully removed pairs
gives information on the distribution of edges (the fewer removed pairs, the more uniform the
distribution). As well as the number of removed pairs, we also record the percentage with respect to
all possible pairs, and the minimum and maximum values on a per-variable basis.

@@

Table~\ref{tab:res} shows the performance of our algorithm selection approach, compared to two
baselines, on the reduced set of 2336 instances. The virtual best solver is the oracle predictor
that, for each instance, chooses the best solver from our portfolio. This is the upper bound of what
an algorithm selection approach can achieve. The single best solver is the one solver from the
portfolio that has the overall best performance across the entire set of instances, at the CPU time
limit of \SI{e8}{\ms}, i.e.\ \GlasgowTwo. We consider it a lower bound on the performance of the
algorithm selection approach.  We are able to solve 34 more instances than the single best solver
within the timeout, with only an additional 13 to the virtual best. In terms of average performance,
we are able to close almost \SI{70}{\percent} of the gap between the single best and the virtual
best solver.

@@

\begin{table}[p]
    \centering\setlength{\tabcolsep}{1em}
\begin{tabular}{lrrr}
  \toprule
model & mean MCP & solved instances & mean performance\\
  \midrule
virtual best & 0 & 2219 & 5822809\\
\LLAMA & 610835 & 2206 & 6435301\\
\GlasgowTwo & 1965158 & 2172 & 7787967\\
   \bottomrule \\
\end{tabular}
\caption{Algorithm selection performance on the reduced set of 2336 instances. MCP is the
misclassification penalty; that is, the additional time required to solve an instance because of
choosing solvers that perform worse than the best. Mean performance is over all instances; when an
instance is not solved, its performance is set to the time limit
(\SI{e8}{\ms}).}\label{tab:res}
\end{table}

@@

Table~\ref{tab:res} shows the performance of the selection model on its own. The performance of the
entire algorithm selection system, including the preprocessing, is shown in Table~\ref{tab:resfull}.
Our system is able to close more than \SI{50}{\percent} of the gap between single and virtual best.

@@

\begin{figure}[p]
    \input{graph-full}
\caption{Number of solved instances over time for the virtual best
solver, \LLAMA, and individual solvers (shown as dotted lines, in the same
colours as Figure~\ref{expTimeGraph}) on the full set of 5725 instances.}
\label{fig:portfolio-ecdf-full}
\end{figure}

@@

The problem of identifying subgraph isomorphisms is a hard computation problem that has many
applications in diverse areas. In this paper, we presented a portfolio of six algorithms from the
literature and two new variants of the \LAD algorithm. We introduced a set of novel features to
characterise subgraph isomorphism problems and leveraged them to select the most appropriate
algorithm from the portfolio for each instance.

We demonstrated that our algorithm selection approach achieves substantial
performance improvements over the single algorithm that has the best performance
on our benchmark set. We showed that combining an algorithm selection approach
with a new incomplete variant of \LAD that is able to detect inconsistencies
and a presolver boosts performance even further.
